---
title: Selection of years from 2010-2019 for calibration/testing
date: "`r Sys.Date()`"
format:
  html:
    code-fold: false
    theme: cosmo
    css: styles.css
    code-tools: 
      source: https://github.com/tbep-tech/tb-box-model/blob/main/docs/hydro_split.qmd
execute:
  echo: true
  warning: false
  message: false
  eval: true
  
lightbox: true
---

This document systematically evaluates groups of years from 2010 to 2019 to determine the optimal split for model calibration/validation.  The optimal split is based on the combination of years where the difference in the total load sum between the year groups is minimized.  Each group has five years each. 

```{r}
library(tidyverse)

load(url('https://github.com/tbep-tech/load-estimates/raw/refs/heads/main/data/mohydat.RData'))

hydat <- mohydat |> 
  filter(year >= 2010 & year <= 2019) |> 
  filter(bay_segment %in% c('Old Tampa Bay', 'Hillsborough Bay', 'Middle Tampa Bay')) |> 
  summarise(
    hy_load = sum(hy_load_106_m3_mo), 
    .by = year
  )
```

Total hydrologic load to Old Tampa Bay, Hillsborough Bay, and Middle Tampa Bay from 2010 to 2019 is shown below.

```{r}
# plot of hydrologic load by year
ggplot(hydat, aes(x = year, y = hy_load)) +
  geom_col(alpha = 0.7) +
  scale_x_continuous(breaks = seq(2010, 2019, 1)) +
  theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  labs(
    x = NULL,
    y = 'Mill m3 / yr',
    title = 'Hydrologic Load by Year',
    subtitle = 'Old Tampa Bay, Hillsborough Bay, Middle Tampa Bay'
  )
```

The unique combinations of years from 2010 to 2019 are generated. 

```{r}
yrs <- hydat$year
lds <- hydat$hy_load

# Generate all possible combinations
combns <- combn(yrs, length(yrs)/2, simplify = FALSE)
```

`r length(combns)` combinations of years are generated.  The absolute difference in total load between each unique combination and its set difference (years not in the unique combination) is calculated.

```{r}
# Calculate absolute difference for each combination
res <- lapply(combns, function(grp1_yrs) {
  
  grp2_yrs <- setdiff(yrs, grp1_yrs)
  
  grp1_sum <- sum(lds[yrs %in% grp1_yrs])
  grp2_sum <- sum(lds[yrs %in% grp2_yrs])
  
  list(
    grp1_yrs = grp1_yrs,
    grp2_yrs = grp2_yrs,
    abs_diff = abs(grp1_sum - grp2_sum)
  )
  
})
```

The optimal grouping is the combination of years with the smallest absolute difference.

```{r}
# Find the combination with smallest mean difference
opt_split <- res[[which.min(sapply(res, `[[`, "abs_diff"))]]
opt_split
```

A plot of the optimal split is shown below.

```{r}
toplo <- hydat |> 
  mutate(
    grp = ifelse(year %in% opt_split$grp1_yrs, 'Group 1', 'Group 2')
  )

# plot of hydrologic load by year
ggplot(toplo, aes(x = year, y = hy_load, fill = grp)) +
  geom_col(alpha = 0.7) +
  scale_x_continuous(breaks = seq(2010, 2019, 1)) +
  theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(), 
    legend.position = 'bottom'
  ) +
  labs(
    x = NULL,
    y = 'Mill m3 / yr',
    fill = 'Split',
    title = 'Hydrologic Load by Year, showing year splits that minimize difference in sums',
    subtitle = 'Old Tampa Bay, Hillsborough Bay, Middle Tampa Bay'
  )
```
